Containers and R---the Rockerverse and beyond

### 2020-01-09 manuscript.Rmd
add rocker-win Reference

[`rocker-win`](https://github.com/nuest/rocker-win) is a proof of concept for running R in Windows-based containers.
It provides selected R versions using three different base images using `microsoft/windowsservercore` and `mcr.microsoft.com/windows/servercore`.
These base images match different Windows versions: Windows Server 2019, Windows Server 2016, and Windows Server, version 1803.
The images are built using two CI services, [Travis CI](https://docs.travis-ci.com/user/reference/windows/) and [Appveyor](https://www.appveyor.com/docs/windows-images-software/), which provide the different Windows versions and therefore support different base images, and published automatically [on Docker Hub](https://hub.docker.com/r/nuest/rocker-win).
The images are built from manually maintained seperate Dockerfiles, are tagged with both Windows-variant and R version, e.g. `nuest/rocker-win:ltsc2019-3.6.2` or `nuest/rocker-win:1803-latest`, and run the `R.exe` process by default.
New R versions are only added on demand.
The explorative project demonstrates the variety and pontentially confusing complexity of Windows on Docker, e.g., because of the different base images, but also provides a starting point for any Windows-constrained user to leverage containerisation, including examples for apps using \pkg{plumber\, \pkg{Shiny}, or R packages with system dependencies such as \pkg{sf}.
The latter is quite comfortable actually, as R ships pre-compiled binaries for Windows.

@@

## Non-Debian Linux images [NN]

- Alpine images
- [Images used by R-Hub](https://github.com/r-hub/rhub-linux-builders) (overlap with CI?)
- https://github.com/jlisic/R-docker-centos
- Fedora > easy to install: https://github.com/ismailsunni/dockeRs

@@

## Interfaces for Docker in R
\label{interfaces}

@@

# inserting a table in the R Journal currently fails
# waiting for this PR to merge
# https://github.com/rstudio/rticles/pull/261
if (packageVersion("rticles") >= "0.12.3") {
  requireNamespace("kableExtra")
  knitr::kable(tab, escape = FALSE) %>%
    kableExtra::row_spec(0, angle = -90)
} else {
  cat("\n\n[Functionality overview table omitted due to a small bug in rticles]\n\n")
}
```

>>>>

[`rocker-win`](https://github.com/nuest/rocker-win) is a proof of concept for running R in Windows-based containers \citep{nust_rocker-win_2019}.
It provides selected R versions using three different base images using `microsoft/windowsservercore` and `mcr.microsoft.com/windows/servercore`.
These base images match different Windows versions: Windows Server 2019, Windows Server 2016, and Windows Server, version 1803.
The images are built using two CI services, [Travis CI](https://docs.travis-ci.com/user/reference/windows/) and [Appveyor](https://www.appveyor.com/docs/windows-images-software/), which provide the different Windows versions and therefore support different base images, and published automatically [on Docker Hub](https://hub.docker.com/r/nuest/rocker-win).
The images are built from manually maintained seperate Dockerfiles, are tagged with both Windows-variant and R version, e.g. `nuest/rocker-win:ltsc2019-3.6.2` or `nuest/rocker-win:1803-latest`, and run the `R.exe` process by default.
New R versions are only added on demand.
The explorative project demonstrates the variety and pontentially confusing complexity of Windows on Docker, e.g., because of the different base images, but also provides a starting point for any Windows-constrained user to leverage containerisation, including examples for apps using \pkg{plumber\, \pkg{Shiny}, or R packages with system dependencies such as \pkg{sf}.
The latter is quite comfortable actually, as R ships pre-compiled binaries for Windows.

@@

## Non-Debian Linux images

- Alpine images
- [Images used by R-Hub](https://github.com/r-hub/rhub-linux-builders) (overlap with CI?)
- https://github.com/jlisic/R-docker-centos

@@

## Interfaces for Docker in R

@@

# inserting a table in the R Journal currently fails
# waiting for new release of articles including this PR to be published: https://github.com/rstudio/rticles/pull/261
# Until then, use remotes::install_github("rstudio/rticles")
if (packageVersion("rticles") >= "0.12.3") {
  requireNamespace("kableExtra")
  knitr::kable(tab, escape = FALSE) %>%
    kableExtra::row_spec(0, angle = -90)
} else {
  cat("\n\n[Functionality overview table omitted due to a small bug in rticles]\n\n")
}
```
